Percolation analysis of transect data

### 2019-06-25 paper.Rmd
updated mit hinweisen von simon

---
title: "Title Goes Here"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

library(percolatransect) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

@@

Einladen der Daten und verbindung:
```{r}
sNL1 <- read_csv("analysis_results/sNL/1m_sNL/analysis_by_radius_sNL1.csv")
sBZ1  <- read_csv("analysis_results/sBZ/sBZ_1m/analysis_by_radius.csv")

sNL1$period  <- "Spätneolihitikum"
sBZ1$period  <-  "Spätbronzezeit"

radius1 <- rbind(sNL1,sBZ1)

@@

```{r anzahl cluster}
library(ggplot2)
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = num_clust, col = period))
```

```{r max größe cluster}
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = max_clust_size, col = period))
```

```{r mean cluster size}
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = mean_clust_size, col = period))
```

```{r anzahl cluster}
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = max_normalized, col = period))
```
# next point: Vergleich mit meinen density daten?

@@

copied from old:
```{r message=FALSE, include=FALSE, cache=FALSE}

@@

```{r anzahl cluster}
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = max_normalized, col = period))+
  geom_density()
```

@@

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

>>>>

---
title: "Percolation analysis of transect data"
author:
  -Schmidt, Sophie C.
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file

abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

#library(percolatransect) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

##

@@

Einladen der Daten und verbindung:
```{r}
#Data cleaning and prep for Sophie

# load data
BefSchnitt <- read.csv2('source_data/Schnittbefunde4.csv', sep = "\t", dec = ",")

# bringing the periods in the right order
BefSchnitt$epoche <- ordered(BefSchnitt$epoche, levels = c("Neolithikum", "Bronzezeit", "Eisenzeit", "undat."))
BefSchnitt$abschnitt <- ordered(BefSchnitt$abschnitt, levels =c('aelter', 'frueh','mittel', 'spaet'))
BefSchnitt$kultur <- ordered(BefSchnitt$kultur, levels=c("Baalberge", "Bernburg", "Kugelamphorenkultur", "Trichterbecher", "Salzmuende", "Schnurkeramik", "Glockenbecher", "Aunjetitz", "Saalemuendungsgruppe", "sBZ/aeEZ", "Hausurnenkultur"))

# subset settlement features
siedlung_BefSchnitt <- subset(BefSchnitt, BefSchnitt$befundart=='Siedlung')

#Subset the settlement features for relevant cultures
# there are 389 features dated as late bronze age and early iron age, that are being analysed in both epochs
siedlung_SnK_Bef <- subset(siedlung_BefSchnitt, siedlung_BefSchnitt$kultur=='Schnurkeramik')
siedlung_SnK_Bef <-siedlung_SnK_Bef[c('PlcIndex', 'Easting', 'Northing')]
write.csv(siedlung_SnK_Bef, "source_data/BefundesNL-minimal.csv", row.names = FALSE)

siedlung_fBZ_Bef <- subset(siedlung_BefSchnitt,
                           siedlung_BefSchnitt$epoche=='Bronzezeit' &
                             siedlung_BefSchnitt$abschnitt=='frueh')
siedlung_fBZ_Bef <- siedlung_fBZ_Bef[c('PlcIndex', 'Easting', 'Northing')]
write.csv(siedlung_fBZ_Bef, "source_data/BefundefBZ-minimal.csv", row.names = FALSE)

siedlung_sBZ_Bef <-subset(siedlung_BefSchnitt, siedlung_BefSchnitt$epoche=='Bronzezeit' & siedlung_BefSchnitt$abschnitt=='spaet')
siedlung_sBZ_Bef <-siedlung_sBZ_Bef[c('PlcIndex', 'Easting', 'Northing')]
write.csv(siedlung_sBZ_Bef, "source_data/BefundesBZ-minimal.csv")

siedlung_sBZ_aeEZ_Bef <- subset(siedlung_BefSchnitt, siedlung_BefSchnitt$epoche=='Eisenzeit' & siedlung_BefSchnitt$abschnitt=='aelter')
siedlung_sBZ_Bef <-siedlung_sBZ_aeEZ_Bef[c('PlcIndex', 'Easting', 'Northing')]
write.csv(siedlung_sBZ_Bef, "source_data/BefundesBZ_aeEZ-minimal.csv")

## run analysis and save results of every run.

### input in source _file.txt
source_file,map_name,shape_file_name,DTM_file_name
BefundesNL-minimal.csv, late neolithic features, alle_Schnitt-Fundstellen.shp,

source_file,map_name,shape_file_name,DTM_file_name
BefundefBZ-minimal.csv, early bronze age features, alle_Schnitt-Fundstellen.shp,

source_file,map_name,shape_file_name,DTM_file_name
BefundesBZ-minimal.csv, late bronze age features, alle_Schnitt-Fundstellen.shp,

source_file,map_name,shape_file_name,DTM_file_name
BefundesBZ_aeEZ-minimal.csv, late bronze age to early iron age features, alle_Schnitt-Fundstellen.shp,

### input in radius_values.txt (stays)

upper_radius,lower_radius,step_value,limit,upper_threshold,lower_threshold,radius_unit
500,2,1, 1000,0,0,1

Außer bei der fBZ: da muss der minimale Radius 4 sein.
## run code (each time new after changing source_file.txt and saving results somewhere)

```{r}
source("create_nodes_list_d.v4.7.R")
source("clustering_script.v3.6.R")
#source("mapping_clusters.v4.12.R")# Fehler in ogrInfo(dsn = dsn, layer = layer, encoding = encoding, use_iconv = use_iconv, : Cannot open data source... weeiiiiird... nochmal auf anderer Maschine testen.

source("cluster_frequency_script.v2.16.R")

Jeweils die Daten in die vorgesehenen Ordnern verschieben.

## für vergleichende plots:

```{r combine all tables}
library(utils)

sNL1 <- read.csv("analysis_results/sNL/1m_sNL/analysis_by_radius.csv")
fBZ1  <- read.csv("analysis_results/fBZ/analysis_by_radius.csv")
sBZ1  <- read.csv("analysis_results/sBZ/sBZ_1m/analysis_by_radius.csv")
aeEZ1  <- read.csv("analysis_results/aeEZ/analysis_by_radius.csv")

sNL1$period  <- "Spätneolihitikum"
fBZ1$period  <- "Frühbronzezeit"
sBZ1$period  <-  "Spätbronzezeit"
aeEZ1$period <- "ältere Eisenzeit"

radius1 <- rbind(sNL1,sBZ1, fBZ1, aeEZ1)

@@

```{r mean cluster size}
library(ggplot2)

ggplot( data = radius1)+
  geom_line(aes(x = radius, y = mean_clust_size, col = period))
```

```{r anzahl cluster}
library(ggplot2)
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = max_normalized, col = period))+
  theme_light()+
  labs(x ="percolation radius", y = "normalized maximum cluster size",
       title = "Comparison of different time periods")

ggsave("normalized_max_clusters.png", height = 10, width =  20, units = "cm", dpi = 300  )

```
# auslesen der "knackpunkte"?

```{r }
## not happy with results here. esp. for EZ.
# google terms: knee of curve, "trade off point"
#https://dataplatform.cloud.ibm.com/analytics/notebooks/54d79c2a-f155-40ec-93ec-ed05b58afa39/view?access_token=6d8ec910cf2a1b3901c721fcb94638563cd646fe14400fecbb76cea6aaae2fb1

#https://www.rdocumentation.org/packages/SamSPECTRAL/versions/1.26.0/topics/kneepointDetection

points <-  c(1,1+which(diff(sNL1$max_normalized)!=0))

sNL1points<- as.data.frame(cbind(sNL1$radius[points], sNL1$max_normalized[points]))

pointssBZ <-  c(1,1+which(diff(sBZ1$max_normalized)!=0))

sBZ1points<- as.data.frame(cbind(sBZ1$radius[pointssBZ], sBZ1$max_normalized[pointssBZ]))

pointssBZ <-  c(1,1+which(diff(sBZ1$max_normalized)!=0))

sBZ1points<- as.data.frame(cbind(sBZ1$radius[pointssBZ], sBZ1$max_normalized[pointssBZ]))

pointsaeEZ <-c(1,1+which(diff(aeEZ1$max_normalized)!=0))
aeEZpoints <-as.data.frame(cbind(aeEZ1$radius[pointsaeEZ],
aeEZ1$max_normalized[pointsaeEZ]))

ggplot()+
  geom_line(data = radius1, aes(x = radius, y = max_normalized, col = period))+
  geom_point(data = sNL1points, aes(x = sNL1points$V1,  y =sNL1points$V2), alpha = 0.5)+
  geom_point(data = sBZ1points, aes(x=sBZ1points$V1, y=sBZ1points$V2), alpha = 0.5)+
  geom_point(data = aeEZpoints, aes(x=aeEZpoints$V1, y=aeEZpoints$V2))

```
Die Points sind bereits wichtige Radii-Angaben, weil der URsprungsdatensatz ja auch einfach bei 1 = 1m anfängt.

Welcher Knickpunkt ist wichtig? Wendepunkte berechnen:

```{r}

# calculate turning points of distance measure vectors

#install.packages("ecp")
library("ecp")

turn_sNL <- e.divisive(as.matrix(diff(sNL1$max_normalized)), k = 1, min.size = nrow(sNL1)/10) #looks for change points, needs matrix, k number of changepoints estimated, min.size = minimum number of observations between change points -> decided to be a tenth of the number of observations so it's not every small change that's documented

sNL_turn_r <-sNL1$radius[turn_sNL$estimates]

# same late bronze age
turn_sBZ <- e.divisive(as.matrix(diff(sBZ1$max_normalized)), k = 1, min.size = nrow(sBZ1)/10)

sBZ_turn_r <-sBZ1$radius[turn_sBZ$estimates]

# same early iron age
aeEZ_turn  <- e.divisive(as.matrix(diff(aeEZ1$max_normalized)), k = 1, min.size = nrow(aeEZ1)/10)

aeEZ_turn_r <- aeEZ1$radius[aeEZ_turn$estimates]

##Tabelle anschauen

```{r}
sNL_turn_r[2]
sBZ_turn_r[2]
aeEZ_turn_r[2]
```

# next point: Vergleich mit meinen density daten?

@@

copied from old:
```{r eval=FALSE, message=FALSE, cache=FALSE, include=FALSE}

@@

```{r anzahl cluster2, eval=FALSE, include=FALSE}
ggplot( data = radius1)+
  geom_line(aes(x = radius, y = max_normalized, col = period))+
  geom_density()
```

@@
